---
title: "Introduction to herdr"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Introduction to herdr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
# Introduction to herdr

**herdr** is an R package designed to calculate **greenhouse gas (GHG) emissions** and **land use** (LU) associated with livestock production.  

Currently supported animals:
- **Cattle**
- **Goat**
- **Sheep**

---

## What the Package Calculates

The package offers specific functions to model the primary sources of environmental impact from livestock based on a detailed **Energy Balance** model. It accounts for energy requirements for maintenance, growth, lactation, pregnancy, work, and wool production.

### Methane (CH₄) Emissions
* **Enteric fermentation:** Emissions produced during the digestive process based on Gross Energy intake and diet digestibility.
* **Manure management:** Emissions derived from Volatile Solids ($VS$) decay, adjusting for the specific conditions (MCF) of diverse management systems (e.g., liquid lagoons, solid storage, pasture, composting).

### Nitrous Oxide (N₂O) Emissions
* **Direct:** Emissions from manure storage/treatment systems.
* **Indirect:** Emissions via nitrogen volatilization ($NH_3$, $NO_x$) and leaching into water systems.

### Land Use
* Calculates the **land area** ($m^2$) required to produce the feed, linking animal intake directly to crop yields.

### Spatial Granularity
* The package is designed to handle multiple **Groups** (e.g., farms, countries) and **Zones** (e.g., climatic regions) simultaneously, allowing for comparative analysis within a single dataset.

---

## 🎯 Calculation Methodology (IPCC 2019)

All emission calculations within herdr adhere to the Tier 2 methods established by the Intergovernmental Panel on Climate Change (IPCC).

Specifically, we use the guidelines from the 2019 Refinement to the 2006 IPCC Guidelines for National Greenhouse Gas Inventories:

*"Volume 4: Agriculture, Forestry and Other Land Use, Chapter 10: EMISSIONS FROM LIVESTOCK AND MANURE MANAGEMENT"*

---

## 🥕 The Role of Diet & Validation

Diet composition is the primary driver of the results. The functions use nutritional data (Energy, Protein, NDF) to determine key parameters like the Methane Conversion Factor ($Y_m$) and Nitrogen Excretion ($N_{ex}$).

**Automatic Validation:**
Before calculating, `herdr` automatically checks consistency between your census (population) and diet files, ensuring every animal defined has a corresponding nutritional profile.

---

## ⬇️ Installation & Loading 

- Install the package directly from GitHub using **remotes**:

```{r install_herdr, eval=FALSE}
# Instalar 'remotes' si es necesario
if (!requireNamespace("remotes", quietly = TRUE)) {
    install.packages("remotes")
}
# Instalar el paquete herdr
remotes::install_github("JuanCBM99/Animal_GEI_LU")
```

- Load the package and explore the default datasets:

```{r load_herdr}
library(herdr)
```
---

## Next Steps

After familiarizing yourself with the package, move on to the Example Data vignette for detailed instructions on:

- Loading and exploring datasets

- Using your own data with templates

- Preparing data for calculations




