---
title: "Running Calculations with AnimalGEILU"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Running Calculations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

---

# Running Calculations with AnimalGEILU
This vignette shows how to use the AnimalGEILU functions to calculate greenhouse gas (GHG) emissions and land use.

We will cover:

* The main summary function (the easiest way to get results).

* How to run individual calculation steps.

* How to filter calculations by animal, type, or zone.

Let's load the package to get started.

```{r load_package}
library(AnimalGEILU)
```

---

## 1. The Easy Way: The Summary Function

*For most uses, you only need one function: resumen_emisiones_land_use().*

This function automatically runs all the individual calculations ($CH_4$, $N_2O$, Land Use) and combines them into a single summary table.

By default, this function also saves a file named Resumen_emisiones_land_use.csv inside an output/ folder in your project.

```{r calculate_all, eval= FALSE}
# Run all calculations for all animals
summary_all <- resumen_emisiones_land_use()

# View the first few rows of the result
head(summary_all)
```

---

## 2. Filtering Your Calculations

You can easily filter the calculations for specific animals, types, or zones. These parameters work for the main summary function and for all individual functions.

* animal: "Cattle", "Sheep", or "Goat"

* type: e.g., "Dairy", "Beef" (specific to "Cattle")

* zone: e.g., "A", "B" (specific to "Cattle" types)

### Example 1: Calculate for Dairy Cattle in Zone A

```{r calculate_example_1, eval= FALSE}
# Run calculations only for Dairy Cattle in Zone A
dairy_A_summary <- resumen_emisiones_land_use(animal = "Cattle", 
                                              type = "Dairy", 
                                              zone = "A")
                                                      
head(dairy_A_summary)
```

Example 2: Calculate for all Sheep

```{r calculate_example_2, eval= FALSE}
# Run calculations for all sheep categories
sheep_summary <- resumen_emisiones_land_use(animal = "Sheep")
        
head(sheep_summary)
```


## 3. Advanced: Running Individual Calculations

You can also run each calculation step individually. 

This is useful if you are only interested in one specific emission source (like enteric methane) or if you want to analyze the intermediate steps.

All functions follow the same pattern: calculate_...()

Methane ($CH_4$) Emissions

Enteric Fermentation:

ch4_enteric <- calculate_emissions_enteric(animal = "Cattle")
head(ch4_enteric)

Manure Management:
ch4_manure <- calculate_CH4_manure(animal = "Cattle")
head(ch4_manure)

Nitrous Oxide ($N_2O$) EmissionsDirect Emissions (from Manure):
n2o_direct <- calculate_N2O_direct_manure(animal = "Cattle")
head(n2o_direct)
Indirect Emissions (from Volatilization):Fragmento de códigon2o_volatilization <- calculate_N2O_indirect_volatilization(animal = "Cattle")
head(n2o_volatilization)
Indirect Emissions (from Leaching):Fragmento de códigon2o_leaching <- calculate_N2O_indirect_leaching(animal = "Cattle")
head(n2o_leaching)
Land UseLand Use for Feed:Fragmento de códigolu_feed <- calculate_land_use(animal = "Cattle")
head(lu_feed)
