---
title: "Example Data in AnimalGEILU"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example Data in AnimalGEILU}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
# Example Data in AnimalGEILU

This package includes several default datasets to get you started immediately.

This guide will show you:

* What data is included by default.
* How to explore that data.
* How to prepare and use your own data for calculations.

Let's load the package to get started:

```{r load_package}
library(AnimalGEILU)
```

---

## 1. Exploring the Included Datasets

The package's data can be grouped into two types: **Input Data** and **Calculation Parameters** (Coefficients and Factors).

💡 **Tip:** You can see a list of *all* available datasets by running `data(package = "AnimalGEILU")`.

### Input Data

This is the primary data describing your animals and their diet. You will likely want to replace these with your own data.

* `categories`: Defines the animal categories.
* `weights`: Contains animal weights (e.g., live weight).
* `diet`: Describes the diet composition for different animals.
* `ingredients`: Nutritional details of the diet ingredients.
* `characteristics`: General diet characteristics.
* 

**Let's look at an example** by exploring the `diet` dataset. First, we load it into our environment, and then we use `head()` to see the first few rows.

```{r explore_diet, eval= FALSE}
# Load the specific dataset
data("diet")

# View the first 6 rows
head(diet)
```

---

## 2. Using Your Own Data

`AnimalGEILU` provides a simple, automatic way to use your custom data.

Instead of you manually loading files (like with `read_csv`), the package will **automatically detect and use your data** if you place it in a specific folder.

Here is the exact workflow:

### The Automatic Workflow

**1. Create the `user_data` Folder**
In the **root directory** of your R project (the same folder where your `.Rproj` file is), create a new folder and name it *exactly* `user_data`.

**2. Get the Templates**
You must use the official CSV templates to ensure your data has the correct format.

* **Download them here:** [github.com/JuanCBM99/Animal_GEI_LU/tree/dev/inst/extdata/templates](https://github.com/JuanCBM99/Animal_GEI_LU/tree/dev/inst/extdata/templates)

Download the templates (e.g., `diet.csv`, `weights.csv`) for the data you want to replace.

**3. Add Your Data and Save**
Open the downloaded CSV templates, fill them with your own project data, and save them inside the `user_data` folder you created in Step 1.

**That's it!**

The next time you run a calculation function from `AnimalGEILU`, the package will automatically see your CSV files inside `user_data` and use them for the calculations instead of the default package data.

---

### 📁 The Golden Rule: Match the Structure

This automatic detection only works if your CSV files **perfectly match the structure of the templates**.

This means you must use the exact same **column names** and provide the correct **data types** (e.g., `numeric` for numbers, `character` for text).

You can always check the required structure of the *default* data (which the templates are based on) by loading it from the package and using `str()`. This is very useful for checking data types.

```{r structure_diet, eval = FALSE}
# This code shows you the structure of the *default* data.
# Your CSV file in the 'user_data' folder must match this structure.

# Load the default 'diet' data from the package
data("diet")

# Check its structure
str(diet)
```

---

## 2. Next Steps

Now that you know how to structure your data, the next vignette will show you how to use the package's calculation functions.

